---
title: "Predicting the surfactant class from 450k arrays"
author: Andreas Mock and Kolja Pocher 
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Introduction

The following tutorial describes the prediction of the"surfactant high" and "surfactant low" classes from methylome data (450k arrays). The predictor requires the `CellMix` and the `repmis` R package.

```{r, message=FALSE, warning=FALSE}
library(CellMix)
library(repmis)
```

The predictor can be loaded by:

```{r}
source_data("https://github.com/andreasmock/surfactantClass/raw/master/surfactantClass.RData")
```


#### Download example data

To illustrate the functionality of the predictor, we download a examplary lung adenocarcinoma sample analysed with the Illumina 450k array from an independent publish study using the Gene Expression Omnibus (Bjaanæs et al., 2016, [GSE66836](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?&acc=GSE66836)). The sample id is GSM1632880.

```{r, eval=TRUE}
sample <- read.table(file="https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?mode=raw&is_datatable=true&acc=GSM1632880&id=21811&db=GeoDb_blob124", sep="\t", header = TRUE)
```

The object `sample` comprises the beta values of 456946 CpGs.

```{r}
dim(sample)
head(sample)
```


#### Predict surfactant class 

The function `surfactantClass` applies a deconvolution algorithm to estimate the fraction of the "surfactant high" and "surfactant low" phenotype. The class label is given according to the phenotype with the highest estimate.

The function requires the vector `all_cpgs` of the CpG labels and a vector or matrix `data` of the corresponding beta values.

```{r}
class_prediction <- surfactantClass(all_cpgs = sample$ID_REF, data = sample$VALUE)
```

```{r}
class_prediction
```

The examplary LUAD sample can be classified to be "surfactant high".

#### Session info

```{r}
sessionInfo()
```

